---
title: "Exploring Hillary Clinton's Emails"
author: "Ben Hamner"
---

So far, [Hillary Clinton has released 7,945 emails](https://www.hillaryclinton.com/hillarys-emails-four-sentences/) in response to a FOIA request. The original emails (released as PDF documents) are available on the [State Department's Website](https://foia.state.gov/Search/Results.aspx?collection=Clinton_Email).

I've transformed this data to two more digestable formats: CSV files and a SQLite database. You can download this transformed data [on Kaggle](https://www.kaggle.com/c/hillary-clinton-emails/data), see the code used to create it [on GitHub](https://github.com/benhamner/hillary-clinton-emails), and share your analytics on [Kaggle Scripts](https://www.kaggle.com/c/hillary-clinton-emails/data).

The transformation process currently has many imperfections, contributions to improve it are welcome!

Let's connect to the database and start exploring it.

```{r include=FALSE}
library.warn <- library
library <- function(package, help, pos = 2, lib.loc = NULL, character.only = FALSE,
  logical.return = FALSE, warn.conflicts = TRUE, quietly = FALSE,
  verbose = getOption("verbose")) {
  if (!character.only) {
    package <- as.character(substitute(package))
  }

  suppressPackageStartupMessages(library.warn(
    package, help, pos, lib.loc, character.only = TRUE,
    logical.return, warn.conflicts, quietly, verbose))}
```

```{r}
library(RSQLite)
db <- dbConnect(dbDriver("SQLite"), "../input/database.sqlite")
```


```{r include=FALSE}
library(xtable)
print.table <- function(table) {
  html <- print(xtable(table), type="html", print.results=FALSE, include.rownames=FALSE)
  cat(paste0("<div style='width:800; overflow:auto; border-width: 2;'><style>td {padding: 3px;} th {padding: 3px;}</style>", html, "</div>"))
}
```

First, let's see what tables we have to work with

```{r results='asis'} 
library(dplyr)
tables <- dbGetQuery(db, "SELECT Name FROM sqlite_master WHERE type='table'")
colnames(tables) <- c("Name")
tables <- tables %>%
          rowwise() %>%
          mutate(RowCount=dbGetQuery(db, paste0("SELECT COUNT(Id) RowCount FROM ", Name))$RowCount[1])
print.table(tables)
```

The emails table contains 7,945 emails that Hillary Clinton sent or received, and the Persons table contains ~500 senders and receivers.

Who sent the most messages?

```{r, fig.height=4, fig.width=8}
commonSenders <- dbGetQuery(db, "
SELECT p.Name, COUNT(p.Name) NumEmailsSent
FROM Emails e
INNER JOIN Persons p ON e.SenderPersonId=p.Id
GROUP BY p.Name
ORDER BY COUNT(p.Name) DESC
LIMIT 10")

library(ggplot2)
ggplot(commonSenders, aes(x=reorder(Name, NumEmailsSent), y=NumEmailsSent)) +
  geom_bar(stat="identity", fill="#53cfff") +
  coord_flip() + 
  theme_light(base_size=16) +
  xlab("") +
  ylab("Number of Emails Sent") + 
  theme(plot.title=element_text(size=14))

commonRecipients <- dbGetQuery(db, "
SELECT p.Name, COUNT(p.Name) NumEmailsReceived
FROM Emails e
INNER JOIN EmailReceivers r ON r.EmailId=e.Id
INNER JOIN Persons p ON r.PersonId=p.Id
GROUP BY p.Name
ORDER BY COUNT(p.Name) DESC
LIMIT 10")

library(ggplot2)
ggplot(commonRecipients, aes(x=reorder(Name, NumEmailsReceived), y=NumEmailsReceived)) +
  geom_bar(stat="identity", fill="#53cfff") +
  coord_flip() + 
  theme_light(base_size=16) +
  xlab("") + 
  ylab("Number of Emails Received") + 
  theme(plot.title=element_text(size=14))
```

Let's look at a sample of some of the messages Hillary Clinton wrote.

```{r results='asis'}
emailsFromHillary <- dbGetQuery(db, "
SELECT p.Name Sender,
       ExtractedBodyText EmailBody
FROM Emails e
INNER JOIN Persons p ON e.SenderPersonId=P.Id
WHERE p.Name='Hillary Clinton'
  AND e.ExtractedBodyText != ''
ORDER BY RANDOM()")
print.table(head(emailsFromHillary))
```

We'll create a helper function that enables us to look at wordclouds of groups of emails.

```{r}
library(tm)
library(wordcloud)
makeWordCloud <- function(documents) {
  corpus = Corpus(VectorSource(tolower(documents)))
  corpus = tm_map(corpus, removePunctuation)
  corpus = tm_map(corpus, removeWords, stopwords("english"))
  
  frequencies = DocumentTermMatrix(corpus)
  word_frequencies = as.data.frame(as.matrix(frequencies))
  
  words <- colnames(word_frequencies)
  freq <- colSums(word_frequencies)
  wordcloud(words, freq,
            min.freq=sort(freq, decreasing=TRUE)[[100]],
            colors=brewer.pal(8, "Dark2"),
            random.color=TRUE)  
}
```

## Here's the common terms in the emails Hillary Clinton sent.
```{r}
makeWordCloud(emailsFromHillary[["EmailBody"]])
```

Let's look at a sample of some of the messages Hillary Clinton received.

```{r results='asis'}
emailsToHillary <- dbGetQuery(db, "
SELECT p.Name Recipient,
       ExtractedBodyText EmailBody
FROM Emails e
INNER JOIN EmailReceivers r ON r.EmailId=e.Id
INNER JOIN Persons p ON r.PersonId=P.Id
WHERE p.Name='Hillary Clinton'
  AND e.ExtractedBodyText != ''
ORDER BY RANDOM()")
print.table(head(emailsToHillary))
```

## Here's the common terms in the emails Hillary Clinton received.
```{r}
makeWordCloud(emailsToHillary[["EmailBody"]][1:2000])
```

## Most common subjects
```{r results="asis"}
commonSubjects <- dbGetQuery(db, "
SELECT MetadataSubject Subject,
       COUNT(Id) NumberOfOccurences
FROM Emails
GROUP BY MetaDataSubject
ORDER BY COUNT(Id) DESC")
print.table(head(commonSubjects, n=20))
```

## Libya Emails
We can pull out emails on relevant topics. For example, here's a sample of emails that are related to Libya.
```{r results='asis'}
libyaEmails <- dbGetQuery(db, "
SELECT p.Name Sender,
       ExtractedBodyText EmailBody
FROM Emails e
INNER JOIN Persons p ON e.SenderPersonId=P.Id
WHERE e.ExtractedBodyText LIKE '%Libya%'
ORDER BY RANDOM()")
print.table(head(libyaEmails))
```
